Probabilistic Inference of Stellar Parameters:\\ Application to Flickering Stars \altaffilmark{\titledag}

### 2015-05-07 version1.0/introduction.tex
hack day cont

Stellar characterization techniaues overview, traditionally frequentist

use of HBMs, particularly for exoplanets lately

Here we introduce an application of HBMs to fundamental stellar chaarcetrization
specifically for flicker

>>>>

% Stellar characterization techniaues overview, traditionally frequentist

% who cares about characterization?

Accurate and precise stellar characterization forms the bedrock of a number
fields within astronomy.
For example, stellar populations, galactic archaeology, binary stars,
asteroseismology and exoplanet studies all rely on inferences of fundamental
stellar parameters.
These stellar parameters are inferred from a variety of observable properties
measured using spectroscopy, photometry, interferometry, etc.
Many stellar parameters are inferred via theoretical stellar models, e.g.
fitting model isochrones to spectroscopic measurements, and
some are inferred using empirically calibrated relations.
Regardless of the approach, be it theoretical or empirical, the methods used
for the inference of stellar parameters are traditionally frequentist.
However, a shift towards probabilistic methods in other areas of astronomy,
and particularly within the exoplanet community, prompts the need to treat
stars in the same probabilistic framework.

The demand for probilistic stellar parameters not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, but also

%A number of recent exoplanet studies have used Hierarchical Bayesian Modeling
%(HBM) to model the dispersion, or intrinsic scatter in relations between
%astrophysical properties.
%For example, \citep{wolfgang:2015} use HBM to recalibrate the mass-radius
%relation for sub-Neptunes and infer the level of astrophysical dispersion

% We use the example of flicker; the brightness fluctuations of stars on
% timescales less than 8 hours, which has been shown to correlate both with
% surface gravity and stellar density, to demonstrate our probilistic approach.

% Previous studies calibrating the relations between flicker, surface
% gravity and stellar density have assumed that the relations between these
% properties are deterministic, i.e. if you know the flicker of a star with
% infinite precision, then you can exactly predict the dependant property.
% We here provide a probabilistic recalibration of the relations between surface
% gravity, stellar density and flicker, allowing for the astrophysical
% dispersion in the relation.
% This probabilitic recalibration will produce predictive {\it distributions}
% for stellar density and surface gravity, given a measurement of flicker.

% use of HBMs, particularly for exoplanets lately

% Here we introduce an application of HBMs to fundamental stellar
% characterization specifically for flicker.

### 2015-05-07 version1.0/manuscript.tex
hack day cont

%\documentclass[12pt,preprint]{aastex}
\documentclass[apjl]{emulateapj}
\usepackage{comment}
\usepackage{ifthen}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{wasysym}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage[%
	colorlinks=true,	% false: boxed links; true: colored links
	urlcolor=blue,		% color of external links
	pdfpagelabels=true,
	hypertexnames=true,
	plainpages=false,
	naturalnames=true,
%	bookmarks=true,         % show bookmarks bar?
%	unicode=false,          % non-Latin characters in Acrobat’s bookmarks
%	pdftoolbar=true,        % show Acrobat’s toolbar?
%	pdfmenubar=true,        % show Acrobat’s menu?
%	pdffitwindow=false,     % window fit to page when opened
%	pdfstartview={FitH},    % fits the width of the page to the window
	pdftitle={Priors for Orbital Eccentricity},    % title
	pdfauthor={David M. Kipping},     % author
%	pdfsubject={Subject},   % subject of the document
%	pdfcreator={Creator},   % creator of the document
%	pdfproducer={Producer}, % producer of the document
%	pdfkeywords={keyword1} {key2} {key3}, % list of keywords
%	pdfnewwindow=true,      % links in new window
	linkcolor=red,          % color of internal links (change box color with linkbordercolor)
	citecolor=green,        % color of links to bibliography
%	filecolor=magenta,      % color of file links
        ]{hyperref}
%

@@

%% Titlepage
\title {Hierarchical Bayesian Inference of Fundamental Stellar Parameters:\\
Calibration of Flicker to Surface Gravity and Mean Density
\altaffilmark{\titledag}}

@@

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Abe et al.}{2000}]{abe:2000} 
Abe, Y., Ohtani, E., Okuchi, T., Righter, K. \& Drake, M., 2000, Origin of the
Earth and Moon, ed. R. M. Canup \& K. Righter (Tucson: Univ. Arizona
Press), 1, 413
\end{thebibliography}

>>>>

%\documentclass[12pt,preprint]{aastex}
\documentclass[apjl]{emulateapj}

\pdfoutput=1

\usepackage{comment}
\usepackage{ifthen}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{wasysym}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage[%
	colorlinks=true,	% false: boxed links; true: colored links
	urlcolor=blue,		% color of external links
	pdfpagelabels=true,
	hypertexnames=true,
	plainpages=false,
	naturalnames=true,
%	bookmarks=true,         % show bookmarks bar?
%	unicode=false,          % non-Latin characters in Acrobat’s bookmarks
%	pdftoolbar=true,        % show Acrobat’s toolbar?
%	pdfmenubar=true,        % show Acrobat’s menu?
%	pdffitwindow=false,     % window fit to page when opened
%	pdfstartview={FitH},    % fits the width of the page to the window
	pdftitle={Priors for Orbital Eccentricity},    % title
	pdfauthor={David M. Kipping},     % author
%	pdfsubject={Subject},   % subject of the document
%	pdfcreator={Creator},   % creator of the document
%	pdfproducer={Producer}, % producer of the document
%	pdfkeywords={keyword1} {key2} {key3}, % list of keywords
%	pdfnewwindow=true,      % links in new window
	linkcolor=red,          % color of internal links (change box color with linkbordercolor)
	citecolor=green,        % color of links to bibliography
%	filecolor=magenta,      % color of file links
        ]{hyperref}
%

@@

%% Titlepage
\title {Probabilistic Inference of Stellar Parameters:\\
Application to Flickering Stars %to Surface Gravity and Mean Density
\altaffilmark{\titledag}}

@@

One of the more recent tools developed to characterize stars is known as 
``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an 
8-hour timescale (dubbed $F_8$) in a broad visible bandpass time series 
photometric light curve, such as that from \textit{Kepler} or the upcoming TESS 
mission. A more detailed account of the proceedure to calculate flicker is 
described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker 
displays a remarkable correlation to the asteroseismically determined parent 
star surface gravities (\logg). Turning this round, the observation implies 
that flicker can be used to blindly infer surface gravities at the level of 
$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light 
curves available through the array of exoplanet transit missions flying and 
scheduled to launch.

\citet{cranmer:2014} demonstrated that stellar surface granulation is expected
to reproduce a flicker effect in close agreement with that observed by 
\citet{bastien:2013}, providing a physically-plausible explanation. Since
surface gravity is highly correlated to mean stellar density (\rhostar) on 
evolutionary tracks, \citet{kipping:2014} showed that flicker can be also 
be used to infer \rhostar, which is more useful for exoplanet transit 
analysis.

Whether one calibrates flicker to \logg\ or \rhostar, there are several aspects
of the problem which are attractive for our purposes of a simple demonstration
of probabilistic inference of stellar parameters. Firstly, in log-log space the
relationship is very simple, appearing to be linear. Secondly, there is a
sufficiently number of points in the sample (439 stars) to constrain a 
population-based model. Thirdly, there is significant excess scatter around the
best-fitting relation implying that a deterministic model is inadequate. For
these reasons, we identify the calibration of flicker to \logg\ and \rhostar\
as a well-posed problem to first demonstrate probabilistic inference in the 
arena of stellar characterization.

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien, 
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014} 
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781, 
124
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014} 
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015} 
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted 
(astro-ph:1504.07557)
\end{thebibliography}
